---
title: "Group 8"
author: "Yilin Chen"
date: "`r format(Sys.Date(), format='%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
setwd("C:/Users/evely/Desktop/UW PhD/02 BIOST 537 Survival Analysis/Hwk/Group 8")
library(tinytex)
library(flexsurv)
library(survival)
library(ComparisonSurv)
library(dplyr)
library(pander)
library(sjPlot)
library(msm)
library(survMisc)
bmt<-read.csv("bmt-2.csv")
```

## Q3
Are any of the measured baseline factors associated with differences in disease-free survival?

```{r Q3, warning=F}
# glimpse
glimpse(bmt)

# time until relapse or death
## by disease group
s.bmt<-survfit(Surv(tdfs,deltadfs)~as.factor(disgroup),data=bmt)
plot(s.bmt)
## by fab
s.bmt<-survfit(Surv(tdfs,deltadfs)~fab,data=bmt)
plot(s.bmt)

# time until death
## by disease group
s.bmt<-survfit(Surv(ts,deltas)~as.factor(disgroup),data=bmt)
plot(s.bmt)
## by fab
s.bmt<-survfit(Surv(ts,deltas)~fab,data=bmt)
plot(s.bmt)

# baseline factors and differences in DFS
## patient age
coxph(Surv(tdfs,deltadfs)~age,data=bmt) # 0.338
## patient gender
coxph(Surv(tdfs,deltadfs)~male,data=bmt) # 0.301
## donor age
coxph(Surv(tdfs,deltadfs)~donorage,data=bmt) # 0.252
## donor gender
coxph(Surv(tdfs,deltadfs)~donormale,data=bmt) # 0.97
## patient CMV
coxph(Surv(tdfs,deltadfs)~cmv,data=bmt) # 0.482
## donor CMV
coxph(Surv(tdfs,deltadfs)~donorcmv,data=bmt) # 0.836
## the wait time from diagnosis to transplantation
coxph(Surv(tdfs,deltadfs)~waittime,data=bmt) # 0.791
## disease group
coxph(Surv(tdfs,deltadfs)~as.factor(disgroup),data=bmt) # 0.0457, 0.1516
## FAB classification 
coxph(Surv(tdfs,deltadfs)~fab,data=bmt) # 0.00427
## prophylactic use of methotrexate
coxph(Surv(tdfs,deltadfs)~mtx,data=bmt) # 0.0944
## hospital center
coxph(Surv(tdfs,deltadfs)~as.factor(hospital),data=bmt) 
```

**Findings**

Disease group and FAB classification are significantly associated with negative outcome.


## Q4
It is generally thought that a GVHD has an anti-leukemic effect. Based on the available data, is occurrence of aGVHD after transplantation associated with improved disease-free survival? Is it associated with a decreased risk of relapse? In view of this, do you consider aGVHD as an important prognostic event?

```{r Q4, warning=F}
# DFS
coxph(Surv(tdfs,deltadfs)~deltaa,data=bmt) # 0.668

## consider time-varying covariate
bmt.tvc=tmerge(data1=bmt,data2=bmt,id=id,deltadfs=event(tdfs,deltadfs),postGVHD=tdc(ta))
head(bmt.tvc)
s.bmt.tvc<-Surv(bmt.tvc$tstart,bmt.tvc$tstop,bmt.tvc$deltadfs)
coxph(s.bmt.tvc~postGVHD,data=bmt.tvc) # 0.413

# relapse
coxph(Surv(tdfs,deltar)~deltaa,data=bmt)  # 0.345

# death
coxph(Surv(ts,deltas)~deltaa,data=bmt) # 0.423

## consider time-varying covariate
bmt.tvc=tmerge(data1=bmt,data2=bmt,id=id,deltas=event(ts,deltas),postGVHD=tdc(ta))
s.bmt.tvc<-Surv(bmt.tvc$tstart,bmt.tvc$tstop,bmt.tvc$deltas) 
coxph(s.bmt.tvc~postGVHD,data=bmt.tvc) # 0.255
```

**Findings**

GVHD and recovery from platelet might be a time-varying covariate?

Not associated with death, risk of relapse, or composite outcome.


## Q5
Among the patients who develop aGVHD, are any of the measured baseline factors associated with
differences in disease-free survival?

```{r Q5, warning=F}
# baseline factors and differences in DFS
## patient age
coxph(Surv(tdfs,deltadfs)~age,data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp()
## patient gender
coxph(Surv(tdfs,deltadfs)~male,data=bmt %>% filter(deltaa==1))  %>% confint() %>% exp()
## donor age
coxph(Surv(tdfs,deltadfs)~donorage,data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp()
## donor gender
coxph(Surv(tdfs,deltadfs)~donormale,data=bmt %>% filter(deltaa==1))  %>% confint() %>% exp()
## patient CMV
coxph(Surv(tdfs,deltadfs)~cmv,data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp()
## donor CMV
coxph(Surv(tdfs,deltadfs)~donorcmv,data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp()
## the wait time from diagnosis to transplantation
coxph(Surv(tdfs,deltadfs)~waittime,data=bmt %>% filter(deltaa==1))  %>% confint() %>% exp()
## disease group
coxph(Surv(tdfs,deltadfs)~as.factor(disgroup),data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp()  # 0.5537, .0463
## FAB classification 
coxph(Surv(tdfs,deltadfs)~fab,data=bmt %>% filter(deltaa==1)) %>% confint() %>% exp() 
## prophylactic use of methotrexate
coxph(Surv(tdfs,deltadfs)~mtx,data=bmt %>% filter(deltaa==1))  %>% confint() %>% exp() 
```

**Findings**

It seems that donor age and disease group are significant.


## Prognostic Model Building

```{r prog, warning=F, message=F}
# disease group + fab
mod1<-coxph(Surv(tdfs,deltadfs)~as.factor(disgroup)+as.factor(fab),data=bmt) 
mod1
# disease group + fab + patient cmv
mod2<-coxph(Surv(tdfs,deltadfs)~cmv+as.factor(disgroup)+as.factor(fab),data=bmt)
mod2
anova(mod1,mod2)
# disease group + fab + donor cmv
mod3<-coxph(Surv(tdfs,deltadfs)~donorcmv+as.factor(disgroup)+as.factor(fab),data=bmt)
anova(mod1,mod3)
# disease group + fab + aGVHD
mod4<-coxph(Surv(tdfs,deltadfs)~deltaa+as.factor(disgroup)+as.factor(fab),data=bmt)
mod4

# include aGVHD as time-varying covariate
bmt.tvc=tmerge(data1=bmt,data2=bmt,id=id,deltadfs=event(tdfs,deltadfs),postGVHD=tdc(ta))
head(bmt.tvc)
s.bmt.tvc<-Surv(bmt.tvc$tstart,bmt.tvc$tstop,bmt.tvc$deltadfs)
mod5<-coxph(s.bmt.tvc~postGVHD+as.factor(disgroup)+as.factor(fab),data=bmt.tvc) # p=0.313
mod5
# further include interaction with donor age
## mod6<-coxph(s.bmt.tvc~postGVHD*donorage+as.factor(disgroup)+as.factor(fab),data=bmt.tvc)
## mod6
## anova(mod5,mod6)

# include mtx
## mod7<-coxph(s.bmt.tvc~postGVHD*donorage+as.factor(disgroup)+as.factor(fab)+deltap,data=bmt.tvc)
## mod7 
## anova(mod6,mod7)

# further include recovery of platelet as time-varying covariate
bmt.tvc=tmerge(data1=bmt,data2=bmt,id=id,deltadfs=event(tdfs,deltadfs),postGVHD=tdc(ta),postPlat=tdc(tp))
head(bmt.tvc)
s.bmt.tvc<-Surv(bmt.tvc$tstart,bmt.tvc$tstop,bmt.tvc$deltadfs)
mod5<-coxph(s.bmt.tvc~postGVHD+postPlat+as.factor(disgroup)+as.factor(fab)+strata(as.factor(hospital)),data=bmt.tvc) # p=0.313
mod5
confint.default(mod5) %>% exp()

# include interaction between donor age and aGVHD
mod6<-coxph(s.bmt.tvc~postGVHD*donorage+postPlat+as.factor(disgroup)+as.factor(fab)+strata(as.factor(hospital)),data=bmt.tvc) # p=0.313
mod6
confint.default(mod6) %>% exp()
anova(mod5,mod6)

# stratifying by hospital center
mod7<-coxph(s.bmt.tvc~postGVHD*donorage+postPlat+as.factor(disgroup)+as.factor(fab)+strata(as.factor(hospital)),data=bmt.tvc)
mod7
anova(mod6,mod7)

# test of PH assumptions
cox.zph(mod5)
cox.zph(mod6)

zp<-cox.zph(mod6,transform=function(time)time)
zp
par(mfrow=c(2,3))
for(i in 1:6){
  plot(zp[i])
  abline(0,0,col="blue",lty=3)
}

## log-log plots for categorical vars
plot(survfit(Surv(tdfs,deltadfs)~as.factor(disgroup),data=bmt),fun="cloglog")
plot(survfit(Surv(tdfs,deltadfs)~as.factor(fab),data=bmt),fun="cloglog")
```

**Findings**

Final model 1: DFS~disease group+FAB+aGVHD (time-varying) + platelet (time-varying)
Final model 2: DFS~disease group+FAB+aGVHD (time-varying)*donor age+ platelet (time-varying)


## Parametric models

```{r aft, warning=F, message=F}
weibullph<-flexsurvreg(Surv(tdfs,deltadfs)~deltaa*donorage+deltap+as.factor(disgroup)+as.factor(fab),data=bmt,dist="weibullph")
weibullph

# compare different model fits
fitweibull<-flexsurvreg(Surv(tdfs,deltadfs)~deltaa*donorage+deltap+as.factor(disgroup)+as.factor(fab),data=bmt,dist="weibull")
fitggamma<-flexsurvreg(Surv(tdfs,deltadfs)~deltaa*donorage+deltap+as.factor(disgroup)+as.factor(fab),data=bmt,dist="gengamma")
fitexp<-flexsurvreg(Surv(tdfs,deltadfs)~deltaa*donorage+deltap+as.factor(disgroup)+as.factor(fab),data=bmt，dist="exp")

plot(survfit(Surv(tdfs,deltadfs)~1, data=bmt), conf.int=FALSE,mark.time=FALSE,
xlim=c(0,1100), xaxt='n', cex.axis=1.5, cex.lab=1.5,
xlab="Time (in days)",ylab="Disease-free survival",lwd=1.5)
axis(1,at=seq(0,1100,by=300),labels=seq(0,1100,by=300),cex.axis=1.5)
lines(fitweibull,col=4,ci=FALSE,lwd=1.8,lty=3)
lines(fitggamma,col=3,ci=FALSE,lwd=1.8,lty=3)
lines(fitexp,col=2,ci=FALSE,lwd=1.8,lty=3)
legend("bottomleft",legend=c("exponential distribution","Weibull distribution",
"generalized gamma distribution",
"nonparametric estimator"),fill=c(2,4,3),cex=0.8)
```



